---
title: Examine topics in embryoid bodies pilot data
author: Peter Carbonetto
output: workflowr::wflow_html
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load packages used in the analysis below, and set the seed.

```{r load-pkgs, message=FALSE}
library(Matrix)
library(dplyr)
library(fastTopics)
library(ggplot2)
library(cowplot)
set.seed(1)
```

This function creates a "hex plot" showing the density of the data
points (topic proportions) as projected onto two principal ccomponents
(PCs).

```{r pca-hex-plot}
pca_hex_plot <-
  function (fit, pcs = c("PC1","PC2"), n = 40, bins = c(0,1,10,100,1000,Inf),
            colors = c("gainsboro","lightskyblue","gold","orange","magenta")) {
  pdat <- as.data.frame(prcomp(fit$L)$x)
  return(ggplot(pdat,aes_string(x = pcs[1],y = pcs[2])) +
         stat_bin_hex(mapping = aes_q(fill = quote(cut(..count..,bins))),
                      bins = n) +
         scale_fill_manual(values = colors) +
         labs(fill = "count") +
         theme_cowplot(font_size = 10))
}
```

Load the $k = 10$ multinomial topic model fit, and compute PCs from
the topic proportions.

```{r load-data}
fit <- readRDS("fit-scd-ex-k=10.rds")$fit
fit <- poisson2multinom(fit)
pca <- prcomp(fit$L)$x
```

TO DO: Describe here what these plots show.

```{r pca-plot-1, fig.height=2, fig.width=8}
p1 <- pca_plot(fit,pcs = c("PC1","PC2"),k = 4) + guides(color = "none")
p2 <- pca_plot(fit,pcs = c("PC1","PC2"),k = 6) + guides(color = "none")
p3 <- pca_hex_plot(fit,pcs = c("PC1","PC2"),bins = c(0,1,10,50,100,Inf))
plot_grid(p1,p2,p3,nrow = 1)
```

*TO DO: Describe here what this code chunk does.*

```{r clustering-1}
n   <- nrow(pca)
x   <- rep("B",n)
pc1 <- pca[,"PC1"]
x[pc1 < -0.05] <- "A"
```

TO DO: Describe here what these plots show.

```{r pca-plot-2, fig.width=7, fig.height=3}
p4 <- pca_plot(fit,pcs = c("PC3","PC4"),k = 8) + guides(color = "none")
p5 <- pca_hex_plot(fit,pcs = c("PC3","PC4"),bins = c(0,1,5,10,100,Inf))
plot_grid(p4,p5,nrow = 1,rel_widths = c(2,3))
```

TO DO: Describe here what this code does.

```{r clustering-2}
pc3 <- pca[,"PC3"]
pc4 <- pca[,"PC4"]
x[pc3 > 2*pc4 + 0.58] <- "C"
clusters <- factor(x)
```

TO DO: Describe here what these plots show.

```{r pca-plot-3, fig.width=7, fig.height=3}
p6 <- pca_plot(fit,pcs = c("PC5","PC6"),k = 1)
p7 <- pca_hex_plot(fit,pcs = c("PC5","PC6"),bins = c(0,1,10,20,100,Inf))
plot_grid(p6,p7,nrow = 1,rel_widths = c(2,3))
```

TO DO: Describe here what these plots show.

```{r pca-plot-4, fig.width=7, fig.height=3}
p8 <- pca_plot(fit,pcs = c("PC7","PC9"),k = 2) + guides(color = "none")
p9 <- pca_hex_plot(fit,pcs = c("PC7","PC9"))
plot_grid(p8,p9,nrow = 1,rel_widths = c(2,3))
```

TO DO: Explain here what these numbers show.

```{r clustering-summary}
table(clusters)
```

TO DO: Describe here what these plots show.

```{r structure-plot, fig.width=7.5, fig.height=2}
rows <- sort(c(sample(which(clusters == "A"),800),
               sample(which(clusters == "B"),800),
               sample(which(clusters == "C"),300)))
p10 <- structure_plot(select(fit,loadings = rows),grouping = clusters[rows],
                      n = Inf,perplexity = 80,gap = 40,num_threads = 4,
                      verbose = FALSE)
print(p10)
```
