---
title: Examine topics in embryoid bodies pilot data
author: Peter Carbonetto
output: workflowr::wflow_html
---

The goal of this analysis is to explore the structure in EB pilot data as
inferred by the multinomial topic model.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load packages used in the analysis below, and set the seed.

```{r load-pkgs, message=FALSE}
library(Matrix)
library(dplyr)
library(fastTopics)
library(ggplot2)
library(cowplot)
set.seed(1)
```

This function creates a "hex plot", showing the density of the data
points (topic proportions) as projected onto two principal ccomponents
(PCs).

```{r pca-hex-plot}
pca_hex_plot <-
  function (fit, pcs = c("PC1","PC2"), n = 40, bins = c(0,1,10,100,1000,Inf),
            colors = c("gainsboro","lightskyblue","gold","orange","magenta")) {
  pdat <- as.data.frame(prcomp(fit$L)$x)
  return(ggplot(pdat,aes_string(x = pcs[1],y = pcs[2])) +
         stat_bin_hex(mapping = aes_q(fill = quote(cut(..count..,bins))),
                      bins = n) +
         scale_fill_manual(values = colors) +
         labs(fill = "count") +
         theme_cowplot(font_size = 10))
}
```

Load the $k = 10$ multinomial topic model fit, and compute PCs from
the topic proportions.

```{r load-data}
fit <- readRDS("fit-scd-ex-k=10.rds")$fit
fit <- poisson2multinom(fit)
pca <- prcomp(fit$L)$x
```

Project the cells onto the first two PCs of the topic proportions.

```{r pca-plot-1, fig.height=2, fig.width=8}
p1 <- pca_plot(fit,pcs = c("PC1","PC2"),k = 4) + guides(fill = "none")
p2 <- pca_plot(fit,pcs = c("PC1","PC2"),k = 6) + guides(fill = "none")
p3 <- pca_hex_plot(fit,pcs = c("PC1","PC2"),bins = c(0,1,10,50,100,Inf))
plot_grid(p1,p2,p3,nrow = 1,rel_widths = c(8,8,11))
```

Judging by this PC projection, the cells can be subdivided, roughly,
into two clusters (noting that there might be other clusters).

```{r clustering-1}
n <- nrow(pca)
x <- rep("B",n)
pc1 <- pca[,"PC1"]
x[pc1 < -0.05] <- "A"
```

In PCs 3 and 4, we see evidence for an additional---but much
smaller---cluster.

```{r pca-plot-2, fig.width=7, fig.height=3}
p4 <- pca_plot(fit,pcs = c("PC3","PC4"),k = 8) + guides(fill = "none")
p5 <- pca_hex_plot(fit,pcs = c("PC3","PC4"),bins = c(0,1,5,10,100,Inf))
plot_grid(p4,p5,nrow = 1,rel_widths = c(8,11))
pc3 <- pca[,"PC3"]
pc4 <- pca[,"PC4"]
x[pc3 > 2*pc4 + 0.58] <- "C"
```

Additional clusters are apparent in the projection onto other PCs.

```{r pca-plot-3, fig.width=6, fig.height=5}
p6 <- pca_plot(fit,pcs = c("PC5","PC6"),k = 1) + guides(fill = "none")
p7 <- pca_hex_plot(fit,pcs = c("PC5","PC6"),bins = c(0,1,10,20,100,Inf))
p8 <- pca_plot(fit,pcs = c("PC7","PC9"),k = 2) + guides(fill = "none")
p9 <- pca_hex_plot(fit,pcs = c("PC7","PC9"))
plot_grid(p6,p7,p8,p9,nrow = 2,rel_widths = c(2,2.75))
```

The cluster that shows up in PCs 7 and 9---corresponding closely to
topic 2---may be picking up some sort of rare cell type.

Using PCA, we can also dig down into individual clusters and look for
subclusters within the clusters. For example, in cluster B, PCs 5 and
6 show interesting substructure:

```{r pca-plot-4, fig.width=5, fig.height=5}
rows <- which(x == "B")
fit2 <- select(fit,loadings = rows)
pca  <- prcomp(fit2$L)$x
n    <- nrow(pca)
y    <- rep("B1",n)
pc3  <- pca[,"PC6"]
y[pc3 > 0.2] <- "B2"
x[rows] <- y
p10 <- pca_plot(fit2,pcs = c("PC5","PC6"),k = c(1,2,8))
print(p10)
```

For example, we can define a subcluster, labeled "B1", that
corresponds closely to topic 2. This substructure does not suggest
clusters that are as distinct as what we saw above, but because topic
2 is picking up a very small subset of cells, it may be useful to
define (somewhat arbitrarily) a cluster so we can analyze it
separately.

```{r clustering-summary}
clusters <- factor(x)
table(clusters)
```

The structure plot summarizes the topic proportions in each of these 4
subsets:

```{r structure-plot, fig.width=7.5, fig.height=2}
rows <- sort(c(sample(which(clusters == "A"),800),
               sample(which(clusters == "B1"),800),
               which(clusters == "B2"),
               sample(which(clusters == "C"),300)))
p11 <- structure_plot(select(fit,loadings = rows),grouping = clusters[rows],
                      n = Inf,perplexity = 80,gap = 40,num_threads = 4,
                      verbose = FALSE)
print(p11)
```

Note that I subsampled the larger subsets so that we can better view
the topics in the smaller subsets.
