---
title: "Exploring a few methods for large-scale multiple linear regression"
author: Peter Carbonetto
output: workflowr::wflow_html
---

*Add overview here.*

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the packages and set the seed so that we may reproduce the
results more easily.

```{r load-pkgs}
library(MASS)
library(SSLASSO)
library(EMVS)
library(ggplot2)
library(cowplot)
set.seed(1)
```

Simulate a data set with correlated variables. This simulation is
similar to Example 1 from [Zou & Hastie 2005][zh2005].

```{r sim-data}
simulate_predictors_decaying_corr <- function (n, p, s = 0.5)
  return(mvrnorm(n,rep(0,p),s^abs(outer(1:p,1:p,"-"))))
simulate_outcomes <- function (X, b, se)
  return(drop(X %*% b - 1 + rnorm(nrow(X),sd = se)))
p      <- 1000
se     <- 3
b      <- rep(0,p)
b[1:3] <- c(3,1.5,2)
Xtrain <- simulate_predictors_decaying_corr(200,p,0.5)
Xtest  <- simulate_predictors_decaying_corr(500,p,0.5)
train  <- list(X = Xtrain,y = simulate_outcomes(Xtrain,b,se))
test   <- list(X = Xtest,y = simulate_outcomes(Xtest,b,se))
btrue  <- b
```

SSLASSO
-------

First let's try both the "adaptive" and "separable" variants of the
[Spike-and-Slab LASSO][sslasso]:

```{r sslasso-1}
sslasso_adapt <-
  with(train,SSLASSO(X,y,penalty = "adaptive",variance = "unknown"))
sslasso_sep <-
  with(train,SSLASSO(X,y,penalty = "separable",variance = "unknown"))
```

Let's take a look at the SSLASSO with separable penalty first. As far
as I can tell, the SSLASSO does not provide an automated way to select
the spike and slab penalties, so I'll choose one by hand:

```{r plot-coef-sslasso-sep, fig.height=2.5, fig.width=2.5}
b <- sslasso_sep$beta[,16]
qplot(btrue,b) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  theme_cowplot(font_size = 12)	 
```

Compare these estimates to the estimates obtained by the adaptive
SSLASSO:

```{r plot-coef-sslasso-adapt, fig.height=2.5, fig.width=2.5}
b <- sslasso_adapt$beta[,16]
qplot(btrue,b) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  theme_cowplot(font_size = 12)	 
```

Another annoying thing about the SSLASSO package is that it does not
provide a "predict" method. So we will compute the predictions by
hand after extracting the :

```{r predict-sslasso, fig.height=2.5, fig.width=5}
b0 <- sslasso_sep$intercept[16]
b  <- sslasso_sep$beta[,16]
y  <- drop(b0 + test$X %*% b)
p1 <- qplot(test$y,y) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  ggtitle(sprintf("separable (cor = %0.3f)",cor(test$y,y))) +
  theme_cowplot(font_size = 12)
b0 <- sslasso_adapt$intercept[16]
b  <- sslasso_adapt$beta[,16]
y  <- drop(b0 + test$X %*% b)
p2 <- qplot(test$y,y) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  ggtitle(sprintf("adaptive (cor = %0.3f)",cor(test$y,y))) +
  theme_cowplot(font_size = 12)
plot_grid(p1,p2)
```

EMVS
----

The [EMVS package][emvs] seems to be better documented (unfortunately,
it was removed from CRAN). It also has two variants with different
priors, the "independent" prior (which is recommended by the authors)
and the "conjugate" prior. Let's compare their performance in this
simulated example.

```{r emvs, results="hide"}
emvs_conj = with(train,
  EMVS(y,X,v0 = seq(0.1,2,length.out = 20),v1 = 10,independent = FALSE))
emvs_ind = with(train,
  EMVS(y,X,v0 = exp(seq(-18,-1,length.out = 20)),v1 = 1,independent = TRUE))
```

*Add text here.*

```{r emvs-conj, fig.height=2.5, fig.width=2.5}
i  <- which.max(emvs_conj$log_g_function)
b  <- with(emvs_conj,betas[i,] * prob_inclusion[i,])
qplot(btrue,b) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  theme_cowplot(font_size = 12)
```

*Add text here.*

```{r emvs-ind, fig.height=2.5, fig.width=2.5}
b <- with(emvs_ind,betas[14,] * prob_inclusion[14,])
qplot(btrue,b) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  theme_cowplot(font_size = 12)
```

*Add text here.*

```{r predict-emvs, fig.height=2.5, fig.width=5}
b0 <- emvs_conj$intersects[i]
b  <- emvs_conj$betas[i,]
y  <- drop(b0 + test$X %*% b)
p1 <- qplot(test$y,y) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  ggtitle(sprintf("conjugate (cor = %0.3f)",cor(test$y,y))) +
  theme_cowplot(font_size = 12)
b0 <- emvs_ind$intersects[14]
b  <- emvs_ind$beta[14,]
y  <- drop(b0 + test$X %*% b)
p2 <- qplot(test$y,y) +
  geom_abline(intercept = 0,slope = 1,color = "deepskyblue",
              linetype = "dashed") +
  ggtitle(sprintf("independent (cor = %0.3f)",cor(test$y,y))) +
  theme_cowplot(font_size = 12)
plot_grid(p1,p2)
```

[emvs]: https://cran.r-project.org/package=EMVS
[sslasso]: https://cran.r-project.org/package=SSLASSO
[zh2005]: https://doi.org/10.1111/j.1467-9868.2005.00503.x
