---
title: Illustration of SuSiE in a "difficult" situation
author: Peter Carbonetto
output: workflowr::wflow_html
---

*Add text here.*

Load packages
-------------

The mvtnorm package is used to simulate the data.

```{r load-pkgs}
library(mvtnorm)
library(varbvs)
library(susieR)
```

Simulate data
-------------

One of the referees was interested in understanding the behaviour of
SuSiE in the following scenario.

There are 5 predictors, and they are all somewhat correlated with each
other. The first two predictors are very correlated with each other,
and the third and fourth predictors are also very strongly
correlated. The fifth predictor is also more correlated with the
others; the weakest correlates are between $(x1,x2)$ and $(x3,x4).

```{r simulate-data}
set.seed(1)
n <- 100
b <- c(1.1,0,-0.75,0,0)
S <- rbind(c(   1, 0.99,  0.5,  0.5, 0.8),
           c(0.99,    1,  0.5,  0.5, 0.8),
           c( 0.5,  0.5,    1, 0.99, 0.8),
           c( 0.5,  0.5, 0.99,    1, 0.8),
           c( 0.8,  0.8,  0.8,  0.8,   1))
X <- rmvnorm(n,sigma = S)
y <- c(X %*% b + 3*rnorm(n))
```

Single-variable regression analysis
-----------------------------------

*Add text here.*

```{r lm} 
for (i in 1:5) {
  dat        <- data.frame(cbind(X[,i],y))
  names(dat) <- c("x","y")
  cat(sprintf("x%d: %0.2e\n",i,
              summary(lm(y ~ x,dat))$coefficients["x","Pr(>|t|)"]))
}
```
