---
title: "Visualizing stationary points of the varbvs objective"
author: "Peter Carbonetto"
date: April 29, 2019
output:
  html_document:
    theme: readable
    include:
      before_body: include/header.html
      after_body: include/footer.html
---

*Add text here.*

```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center")
```

Analysis settings
-----------------

These variables specify the number of samples to simulate ("n"), the
correlation between the two simulated variables, X1 and X2 ("xcor"),
and the regression coefficients used to simulate the outcomes
("beta").

```{r analysis-settings}
n    <- 40
xcor <- 0.99
beta <- c(3,0)
```

These two variables specify how the variational parameters are
initialized in the three varbvs runs.

```{r varbvs-init}
alpha0 <- rbind(c(0,1),
                c(1,0),
                c(1,1))
mu0    <- rbind(c(0,3),
                c(2.5,0),
                c(1,1.5))
```

For creating the surface plot of the variational objective, the
objective is evaluated at all grid points specified by "mu_grid".

```{r mu-grid}
mu_grid <- seq(-1,3.5,0.01)
```

Set up environment
------------------

Load a few packages as well as some functions used in the analysis
below.

```{r load-pkgs, warning=FALSE, message=FALSE}
library(MASS)
library(varbvs)
library(ggplot2)
library(cowplot)
source("visualize_varbvs_surface_functions.R")
```

Initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

Generate data set
-----------------

Simulate the $n \times 2$ matrix of predictors, in which the two
predictors are strongly correlated.

```{r sim-X}
S <- rbind(c(1,xcor),c(xcor,1))
X <- mvrnorm(n,c(0,0),S)
X <- scale(X,center = TRUE,scale = FALSE)
```

Simulate the continuously-valued outcomes.

```{r sim-y}
y <- drop(X %*% beta + rnorm(n))
y <- y - mean(y)
```

Fit model
---------

Run the co-ordinate ascent updates with three different initial
estimates of the variational parameters.

```{r fit-model}
fit1 <- varbvs(X,NULL,y,sigma = 1,sa = 1,logodds = 0,tol = 0,
               alpha = alpha0[1,],mu = mu0[1,],verbose = FALSE)
fit2 <- varbvs(X,NULL,y,sigma = 1,sa = 1,logodds = 0,tol = 0,
               alpha = alpha0[2,],mu = mu0[2,],verbose = FALSE)
fit3 <- varbvs(X,NULL,y,sigma = 1,sa = 1,logodds = 0,tol = 0,
               alpha = alpha0[3,],mu = mu0[3,],verbose = FALSE)
pts <- as.data.frame(rbind(fit1$beta,fit2$beta,fit3$beta))
print(pts)
```

The co-ordinate ascent updates should converge to three different
stationary points of the K-L divergence (equivalently, the variational
lower bound).

Plot variational objective surface
----------------------------------

Compute the value of the variational objective (the K-L divergence) at
each grid point $(\mu_1, \mu_2)$, where $\mu_j$ gives the
(approximate) posterior mean of the jth regression coefficient given
that the coefficient is nonzero.

```{r compute-surface}
pdat <- expand.grid(mu1 = mu_grid,mu2 = mu_grid)
pdat <- compute_kl_grid(X,y,pdat)
```

Create a contour plot to visualize the 2-d surface of the
Kullback-Divergence divergence with respect to the variational
estimates of the posterior mean coefficients.

```{r create-contour-plot, fig.width=6, fig.height=5.5}
p1 <- create_contour_plot(pdat,pts)
print(p1)
```

This contour plot confirms that there are indeed three stationary
points (local minima) of the variational objective: at two of these
stationary points, one of the regression coefficients is zero, whereas
both coefficients are nonzero, and nearly equal, at the third
stationary point.

Session info
------------

This is the version of R and the packages that were used to generate
these results.

```{r session-info}
sessionInfo()
```
